---
title: "Introduction to rebrown through Examples"
author: Yingfa Xie
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_depth: 2
    toc_float: true
bibliography: ../inst/bib/rebrown.bib
vignette: >
  %\VignetteIndexEntry{Introduction to rebrown through Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In this vignettes, we briefly introduce how to explore recurrent event data by 
first hitting time model of reflected Brownian motion (RBM) with the **rebrown** 
package. The **rebrown** package implements the density, distribution, quantile, 
and random number generation function of the first hitting time (FHT) distribution 
(hitting down case) of RBM [@hu2012hitting]. The function `simData()` 
in **rebrown** can be used for simulating recurrent event (gap) times. 
Fitting recurrent data set with FHT model is implemented with `fitFHTRBM()`. 
The sample paths of the reflected Brownian motion can be simulated with 
`simPath()` function, which helps to illustrate when an event occur.


# Introduction 

## First hitting time density and distribution function 
Consider a no-drift Brownian motion $X(t)$ with volatility $\sigma$, 
upper reflection barrier $\kappa > 0$, lower barrier $\nu$, and starting point 
$X(0) = x_0 \in [\nu, \kappa]$. The first time when $X(t)$ hits $\nu$ is 
$\tau := \inf \{t > 0; X(t) = \nu \}$. 
The density and distribution function of $\tau$ are, respectively [@hu2012hitting],
\begin{align}
  f(t|x_0, \nu, \kappa, \sigma) = \sum_{n=1}^{\infty} c_n \lambda_n e^{-\lambda_n t}, 
   \quad F(t|x_0, \nu, \kappa, \sigma) = 1 - \sum_{n=1}^{\infty} c_n e^{-\lambda_n t}, 
\end{align}
where $t > 0$, and for $n = 1, 2, \ldots$,
\begin{align*}
  \lambda_n= \frac{(2n-1)^2 \sigma^2 \pi^2 }{8(\kappa - \nu)^2},
  \mbox{ and }
  c_n       = \frac{(-1)^{n + 1}4 }{(2n-1)\pi}
                \cos \left( \frac{ (2n - 1)\pi(\kappa - x_0)}{2(\kappa - \nu)} \right).
\end{align*}
Note that $0 < \lambda_1 < \lambda_2 < \cdots$, $\lambda_n \to \infty$, and
$\sum_{n=1}^{\infty} c_n = 1$.


Random number generation with rejection sampling (the tail can be
shown to be bounded by that of an exponential distribution).


## First hitting time model for recurrent events
The recurrent events can be modeled based on a sequence of reflected Brownian 
motions hitting a lower barrier $\nu$. The reflected Brownian motion restarts 
at a known starting point $x_0$ after hitting the lower barrier $\nu$. 
The gap times between two successive events follow the FHT distribution. 
The starting point $x_0$ and lower barrier $\nu$ are set to be fixed and assumed 
known.


To incorporate covariate and heterogeneity, the model 
for $\sigma_i$ and $\kappa_i$ of subject $i$ with log links are
\begin{equation}
  \log(\sigma_i) = X_i^{\top} \boldsymbol{\beta} + Z_{i1}, \\
  \log(\kappa_i - x_0) = X_i^{\top} \boldsymbol{\alpha} + \gamma Z_{i1} + Z_{i2}',
\end{equation}
where $X_i$ is a $p$-dimensional covariate vector,
$\boldsymbol{\beta}$ and $\boldsymbol{\alpha}$ are $p$-dimensional 
regression coefficient vectors, and $Z_{i1}$ and $Z_{i2}'$ are 
normal random effects with mean zero,
variance $\theta_1$ and $\theta_2'$, respectively.


## Getting started
```{r, message=FALSE}
library(rebrown)
set.seed(1)
```


## Simulate sample path 
The sample paths help to illustrate when the event occur. `simPath()` requires 
user to pass a vector of time index and the four parameters of the reflected 
Brownian motion. The sample paths below illustrate $2$ events occurred and 
the $3$th is censored during $200$ units of follow-up time. 
```{r}
time_idx <- seq(0, 200, 1)
sim_list <- simPath(x0 = 10, nu = 3.9, kappa = 20, sigma = 1.5,
                    time_idx, plot_path = TRUE)
```


## Simulate event (gap) times 
The `simData()` is used for simulating event (gap) time for subject given the
covariate. For simulating from the full model, one can assign value for the 
regression coefficients in $\sigma$ and $\kappa$, 
the coefficient of the frailty $\gamma$, 
and the frailty variance $\boldsymbol{\theta}$. One can simulate the event time
from independent frailty model by assigning $\gamma = 0$. Furthermore, if the 
same frailty term is attempted to set in both $\sigma$ and $\kappa$, one can 
assign the second element of $\boldsymbol{\theta}$ to be $0$. The simulated data 
set records the calender time of the events for each subject.


```{r}
size <- 200
x1 <- rnorm(size)                
x2 <- rpois(size, 2) + 0.5 * x1 + runif(size)
X <- data.frame("x1" = x1,
                "x2" = x2)
X <- scale(X)
# simulate event time data set from independent frailty model
simuData <- simData(size = size, 
                    endTime = round(runif(size, 160, 180)),
                    X = cbind("Intercept" = rep(1, size), X),
                    sigmaCoef = c(1, 0.2, -0.2),
                    kappaCoef = c(3, 0.1, 0.3),
                    gamma = 0, 
                    theta = c(0.2, 0.3),
                    x0 = 10, nu = 3.9)
```


The simulated data contains `time`, `id`, `status`, and covariate. 
```{r}
head(simuData, 10)
```


## Fit the first hitting time model 
To fit the data with FHT model, one can define the response by creating a `Recur`
object implemented by **R** pacakge **reda** [@reda-package]. 
The formula for the FHT model is given as 
`Recur(time = time, id = id, event = event) ~ x1 + x2|x1 + x2`. 
The left hand side of the vertical bar is given as the covariates incorporated 
into volatility $\sigma$ while the right hand side is given as the covariates 
incorporated into upper reflection barrier $\kappa$. The frailty argument 
can be set to be one of the three types, `Correlated`, `Independent`, or 
`StrgCorrelated`. The function `fitFHTRBM` implements the MCMC with
R package nimble and returns MCMC chain(s) for the model parameters. 
The number of thinning, the burn-in period, and the number of iteration of 
the MCMC chain can be set up by the user. 
```{r, eval = FALSE}
mcmc <- fitFHTRBM(Recur(time = time, id = id, event = event) ~ x1 + x2 | x1 + x2,
                  data = simuData[, -4], # remove intercept column
                  frailty = "Correlated",
                  thin = 10, nburnin = 2000, niter = 10000, nchains = 1)
```


# Reference

